\bf\mftitle

### 2012-03-30 doc/jInfer/architecture-trunk/architecture.tex
Architecture documentation renamed to architecture-1.0. It was also copied and modified to create architecture-trunk documenting architecture in trunk.

The process by which jInfer infers the resulting schema from various inputs (inference process) is summarized by fig. \ref{inference_process}. From the high-level viewpoint, it consists of three consecutive steps carried out by three different modules:

@@

	\item Initial Grammar (IG) generation: done by the \jmodule{Initial Grammar Generator (IGG)} module, this is the process of converting all of the inputs to IG representation. All documents, schemas and queries selected as input are evaluated, simple rules are extracted and in the end sent to the next step.\\

@@

	\item Simplification: done by the \jmodule{Simplifier} module, this is the process of simplifying, compressing or somehow compactly describing the IG by a smaller number of (more complex) rules (exactly one rule for each element). User interaction might be used in this step to help achieve better simplification. At the end of this step, all rules are sent to the export step.\\

@@

@@

	For element person, even when simplified grammar specifies its occurence to at least once, at most 3 times, as DTD has no such construct, export module have to do some magic. Situation is even worse for elements that contain simple data inside simplified rule. Only way to express mixed content in DTD is to use \verb.(#PCDATA | note)*. construct. Even if regular expression is complicated, export module has to do this ``flattening''.texmaker

@@

\subsection{Programmatic view}
From developer's point of view, inference modules are just properly annotated classes (see \ref{section_lookups}) implementing one of the following interfaces (all found under \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference})
\begin{itemize}
	\item \code{IGGenerator}
	\item \code{Simplifier}
	\item \code{SchemaGenerator}
\end{itemize}
A nice way to name such a class is by adding \code{-Impl} to the name of implemented interface, for example \code{Sim\.pli\.fier\.Impl}.
Annotation required for the framework to recognize such a class as an inference module is the following
\begin{verbatim}
	@ServiceProvider(service = <interface>.class)
\end{verbatim}
for example
\begin{verbatim}
	@ServiceProvider(service = Simplifier.class)
\end{verbatim}
The most important method in each module is \code{start}, defined in each of the interfaces. This method is called by the framework when the respective step of inference is to be executed. It has always two parameters: the actual input data for the module, and a callback object to report to when this step is finished. We will look at both parameters now in more detail.

@@

@@

\subsubsection{Module output}
Second parameter of each inference module's \code{start} method is a callback object. There are 3 callback interfaces defined in the \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference} package
\begin{itemize}
	\item \code{IGGeneratorCallback}
	\item \code{SimplifierCallback}
	\item \code{SchemaGeneratorCallback}
\end{itemize}
Each callback interface naturally belongs to the similarly named module interface. As their respective interfaces, also callbacks define one crucial method: \emph{finished}. Each inference module is responsible for invoking this method on the callback it got as a parameter, after it has finished its work and has results to be passed on. Again, these 3 finished methods have different arguments based on the inference module.

\code{IGGeneratorCallback.finished()} and \code{SimplifierCallback.finished()} have a grammar (Initial Grammar in case of IGG) as their only argument.

@@

A new instance of \code{Runner} is constructed for each inference run. While being created, \code{Runner}  loads the preferences for current project and looks up user-selected inference modules. Also, callback objects pointing back to methods in \code{Runner} are created. The inference process itself is then as follows
\begin{figure}
\centering\includegraphics[scale=\myscale]{runner}
\caption{Runner} \label{runner}
\end{figure}
\begin{enumerate}
	\item Selected IGG's \code{start} is encapsulated with error/interruption handling and executed, passing \code{Input} and first callback as parameters.
	\item When IGG finishes, it invokes callback's \code{finish} method, passing the IG as parameter.
	\item This in turn causes \code{Runner} to encapsulate and execute Simplifier's \code{start}, passing IG from the first callback and the second callback as parameters.
	\item When Simplifier finishes, it invokes callback's \code{finish} method, passing the simplified grammar as parameter.
	\item This again causes \code{Runner} to encapsulate and execute SchemaGen's \code{start}, passing the simplified grammar from the second callback and the third callback as parameters.
	\item SchemaGen finishes and invokes the last callback's \code{finish}, passing the resulting schema and its extension as parameters.
	\item \code{Runner} receives the resulting schema and based on preferences, saves it to a file, displays it, etc.
\end{enumerate}

@@

\noindent Inference modules:
\begin{itemize}
	\item \jmodule{Runner}: as described in the section \ref{section_runner}, this module encapsulates the inference process (including the fact that inference consists of 3 phases as described before). Most important class: \code{Runner}.
	\item \jmodule{BasicIGG}: this module contains an extensible implementation of \jmodule{IGGenerator}. Basic version handles all XML-typed input documents, DTD schemas and XPath queries. To find out more about the extensibility to handle other outputs, refer to \jmodule{BasicIGG}'s documentation.
	\item \jmodule{XSDImport}: demonstrating \jmodule{BasicIGG}'s extensibility, this is a base module for XSD input.
	\item \jmodule{XSDImportSAX}: XSD input handling using SAX parser.
	\item \jmodule{XSDImportDOM}: XSD input handling using DOM parser.
	\item \jmodule{TwoStepSimplifier}: a \jmodule{Simplifier} implementation. Implements \cite{ahonen} merging state algorithm with basic handling of attributes. Read the separated documentation of module.
	\item \jmodule{BasicDTDExporter}: this is a \jmodule{SchemaGenerator} implementation with output to DTD.
	\item \jmodule{BasicXSDExporter}: this is a \jmodule{SchemaGenerator} implementation with output to XSD.
\end{itemize}

\noindent Utility modules:
\begin{itemize}
	\item \jmodule{BasicRuleDisplayer}: module capable of displaying a grammar in a graphical way.
	\item \jmodule{TreeRuleDisplayer}: more advanced version of the former using \jmodule{JUNG}'s graphing capabilities.
	\item \jmodule{AutoEditor}: encapsulation of an interactive finite state automaton editor.
	\item \jmodule{JUNG}: wrapper module for the JUNG (see \cite{jung}).
	\item \jmodule{Options}: NBP specific: this module adds a jInfer category in NetBeans Options window.
	\item \jmodule{ProjectType}: NBP specific: this module defines a jInfer project type, along with its input and output files, settings etc.
	\item \jmodule{XPathFileType}: NBP specific: this module teaches NetBeans to recognize a \code{.xpath} file extension.
\end{itemize}
The dependencies between all the modules are summarized in figure \ref{module_deps}. Figure \ref{modules_inference} shows modules implementing the 3 main inference interfaces described in \ref{section_inference_process}.

@@

jInfer uses tokens in the following way: \jmodule{Runner}, as the module encapsulating the inference process, and \jmodule{ProjectType} as an encapsulation of jInfer projects both require the following 3 tokens.
\begin{itemize}
	\item \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.IGGenerator}
	\item \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.SchemaGenerator}
	\item \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.Simplifier}
\end{itemize}
All inference modules declare their respective token as provided. This way it is ensured that jInfer will run only if there is at least one implementation of each of the inference modules, regardless of the implementation itself.

>>>>

The process by which jInfer infers the resulting schema from various inputs (inference process) is summarized by fig. \ref{inference_process}. From the high-level viewpoint, it consists of four consecutive steps carried out by four different modules:

@@

	\item Initial Grammar (IG) generation: done by the \jmodule{Initial Grammar Generator (IGG)} module, this is the process of converting of the inputs to IG representation. Documents, schemas and queries selected as input are evaluated, simple rules are extracted and in the end sent to the next step. Not all types of inputs have to be used to create the initial grammar, as will be described in step 3.\\

@@

	\item Simplification: done by the \jmodule{Simplifier} module, this is the process of simplifying, compressing or somehow compactly describing the IG by a smaller number of (more complex) rules (exactly one rule for each element). User interaction might be used in this step to help achieve better simplification. At the end of this step, all rules are sent to the next step.\\

@@

	\item Non-grammatical input processing: done by the \jmodule{Non-grammatical Input Processor}, in this step, certain inputs are exploited to extract any non-grammatical information that can help in the inference. The information is then merged into the grammar which is sent to the last step. Inputs processed in this step can be inputs that were used to create the initial grammar, and, as well inputs that were not touched by the \jmodule{IGG} module. This step is not a crucial part of the inference, it's purpose is to provide a possibility to easily extend the inference process by incorporating other non-grammatical inputs. Therefore, the inputs and the way of merging the extracted information with the grammar are not specified in detail and they depend on an implementation.

@@

	For element person, even when simplified grammar specifies its occurrence to at least once, at most 3 times, as DTD has no such construct, export module have to do some magic. Situation is even worse for elements that contain simple data inside simplified rule. Only way to express mixed content in DTD is to use \verb.(#PCDATA | note)*. construct. Even if regular expression is complicated, export module has to do this ``flattening''.texmaker

@@

\subsection{Programmatic view}
From developer's point of view, inference modules are just properly annotated classes (see \ref{section_lookups}) implementing one of the following interfaces (all found under \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference})
\begin{itemize}
	\item \code{IGGenerator}
	\item \code{Simplifier}
	\item \code{NonGrammaticalInputProcessor}
	\item \code{SchemaGenerator}
\end{itemize}
A nice way to name such a class is by adding \code{-Impl} to the name of implemented interface, for example \code{Sim\.pli\.fier\.Impl}.
Annotation required for the framework to recognize such a class as an inference module is the following
\begin{verbatim}
	@ServiceProvider(service = <interface>.class)
\end{verbatim}
for example
\begin{verbatim}
	@ServiceProvider(service = Simplifier.class)
\end{verbatim}
The most important method in each module is \code{start}, defined in each of the interfaces. This method is called by the framework when the respective step of inference is to be executed. It has always two parameters: the actual input data for the module, and a callback object to report to when this step is finished. We will look at both parameters now in more detail.

@@

\jmodule{Non-grammatical Input Processor} takes both grammar and \code{Input} object.

@@

\subsubsection{Module output}
Second parameter of each inference module's \code{start} method is a callback object. There are 3 callback interfaces defined in the \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference} package
\begin{itemize}
	\item \code{IGGeneratorCallback}
	\item \code{SimplifierCallback}
	\item \code{NonGrammaticalInputProcessor}
	\item \code{SchemaGeneratorCallback}
\end{itemize}
Each callback interface naturally belongs to the similarly named module interface. As their respective interfaces, also callbacks define one crucial method: \emph{finished}. Each inference module is responsible for invoking this method on the callback it got as a parameter, after it has finished its work and has results to be passed on. Again, these 3 finished methods have different arguments based on the inference module.

\code{IGGeneratorCallback.finished()} and \code{SimplifierCallback.finished()} have a grammar (Initial Grammar in case of IGG) as their only argument. \code{NonGrammaticalInputProcessorCallback.finished()} has that argument, as well.

@@

A new instance of \code{Runner} is constructed for each inference run. While being created, \code{Runner}  loads the preferences for current project and looks up user-selected inference modules. Also, callback objects pointing back to methods in \code{Runner} are created. The inference process itself is then as follows
\begin{figure}
\centering\includegraphics[scale=\myscale]{runner}
\caption{Runner} \label{runner}
\end{figure}
\begin{enumerate}
	\item Selected IGG's \code{start} is encapsulated with error/interruption handling and executed, passing \code{Input} and first callback as parameters.
	\item When IGG finishes, it invokes callback's \code{finish} method, passing the IG as parameter.
	\item This in turn causes \code{Runner} to encapsulate and execute Simplifier's \code{start}, passing IG from the first callback and the second callback as parameters.
	\item When Simplifier finishes, it invokes callback's \code{finish} method, passing the simplified grammar as parameter.
	\item This again causes \code{Runner} to encapsulate and execute NonGrammaticalInputProcessor's \code{start}, passing the simplified from the second callback and the third callback as parameters.
	\item When NonGrammaticalInputProcessor finishes, it invokes callback's \code{finish} method, passing the merged simplified grammar as parameter.
	\item This once again causes \code{Runner} to encapsulate and execute SchemaGen's \code{start}, passing the grammar from the third callback and the fourth callback as parameters.
	\item SchemaGen finishes and invokes the last callback's \code{finish}, passing the resulting schema and its extension as parameters.
	\item \code{Runner} receives the resulting schema and based on preferences, saves it to a file, displays it, etc.
\end{enumerate}

@@

\noindent Inference modules:
\begin{itemize}
	\item \jmodule{Runner}: as described in the section \ref{section_runner}, this module encapsulates the inference process (including the fact that inference consists of 4 phases as described before). Most important class: \code{Runner}.
	\item \jmodule{BasicIGG}: this module contains an extensible implementation of \jmodule{IGGenerator}. Basic version handles all XML-typed input documents, DTD schemas and XPath queries. To find out more about the extensibility to handle other outputs, refer to \jmodule{BasicIGG}'s documentation.
	\item \jmodule{XSDImport}: demonstrating \jmodule{BasicIGG}'s extensibility, this is a base module for XSD input.
	\item \jmodule{XSDImportSAX}: XSD input handling using SAX parser.
	\item \jmodule{XSDImportDOM}: XSD input handling using DOM parser.
	\item \jmodule{XQueryImporter}: XQuery input handling incorporating lexical and syntax analyses as proposed in \cite{mikula}.
	\item \jmodule{TwoStepSimplifier}: a \jmodule{Simplifier} implementation. Implements \cite{ahonen} merging state algorithm with basic handling of attributes. Read the separated documentation of module.
	\item \jmodule{BasicXQueryProcessor}: a \jmodule{Non-grammatical Input Processor} implementation. Implements algorithm proposed in \cite{mikula} processing XQuery queries as an non-grammatical input and refining the inference process by information extracted from the queries.	
	\item \jmodule{BasicDTDExporter}: this is a \jmodule{SchemaGenerator} implementation with output to DTD.
	\item \jmodule{BasicXSDExporter}: this is a \jmodule{SchemaGenerator} implementation with output to XSD.
\end{itemize}

\noindent Utility modules:
\begin{itemize}
	\item \jmodule{BasicRuleDisplayer}: module capable of displaying a grammar in a graphical way.
	\item \jmodule{TreeRuleDisplayer}: more advanced version of the former using \jmodule{JUNG}'s graphing capabilities.
	\item \jmodule{AutoEditor}: encapsulation of an interactive finite state automaton editor.
	\item \jmodule{JUNG}: wrapper module for the JUNG (see \cite{jung}).
	\item \jmodule{Options}: NBP specific: this module adds a jInfer category in NetBeans Options window.
	\item \jmodule{ProjectType}: NBP specific: this module defines a jInfer project type, along with its input and output files, settings etc.
	\item \jmodule{XPathFileType}: NBP specific: this module teaches NetBeans to recognize a \code{.xpath} file extension.
\end{itemize}
The dependencies between all the modules are summarized in figure \ref{module_deps}. Figure \ref{modules_inference} shows modules implementing the 4 main inference interfaces described in \ref{section_inference_process}.

@@

jInfer uses tokens in the following way: \jmodule{Runner}, as the module encapsulating the inference process, and \jmodule{ProjectType} as an encapsulation of jInfer projects both require the following 4 tokens.
\begin{itemize}
	\item \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.IGGenerator}
	\item \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.SchemaGenerator}
	\item \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.NonGrammaticalInputProcessor}
	\item \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.Simplifier}
\end{itemize}
All inference modules declare their respective token as provided. This way it is ensured that jInfer will run only if there is at least one implementation of each of the inference modules, regardless of the implementation itself.
